Delayed effects of climate on vital rates lead to demographic divergence in Amazonian forest fragments

### 2021-08-24 review/Response to reviews.Rmd

<!--# make a table I guess? -->

##  {#R1-4}

> Related to above, I think it's important to see more of the raw data than are presented in Fig.
> 1.
> (I'm thinking of a version of Fig 3 that shows size in t+1, survival, and flowering rather than effect size on the y-axis, with the data and lines showing the shape of the GAMs).
> This could be an appendix, but it would help with interpreting Figures 3 -- 6.

We changed Fig 3 to be on the response scale as suggested (see our response to \@ref(R1-figs), but showing the raw data made the plots visually cluttered, in our opinion.
Another way we approached this comment is by excluding areas of the heatmaps in figures 4-6 that are far from observed data, equivalent to not drawing regression lines beyond the range of data.
We implemented this with the `exclude.too.far()` function in the `mgcv` package.
Displaying raw data on the heatmaps was too messy and not informative in a 2d plot.

@@

##  {#R1-5}

> I suggest that a little more about the limitations of this modeling approach in the discussion would help set the context for interpreting the results and make the paper more robust.

<!--# The following isn't in the main text yet.  Just ideas. -->

The main limitation for detecting lagged effects with any method is data, and the specifics of data limitations have not yet been systematically explored.
@tellerLinkingDemographyDrivers2016 used simulations to show that 20--25 years of data were required to detect lagged effects using their method, but this cutoff depends on the effect size and the number of climate extremes in the dataset.
Additionally, there is some evidence that appropriate spatial replication could drastically reduce the number of years of data required, but this is unpublished currently <!--# cite Aldo Compagnoni's ESA talk -->.

> In some ways it seems like the analysis is adding more precision (by using monthly climate data) than the plants perceive and that would be measured in a yearly census.

On the contrary, averaging weather conditions over an entire year likely obscures ways plants perceive environmental variation.
For example, a drought in the wet season may have a different effect than a drought in the dry season.
This differential response to season- or month-level climate data is observable in yearly plant census.
Collapsing finer temporal scale environmental drivers to yearly averages (or a few means of arbitrarily chosen climate windows) is exactly what we avoid with this method.

> While the authors mention this, what about adding something short about when using this type of analysis is critical and when it might not be (should all demographers adopt this approach? Why or why not?)

We have added additional detail to the conclusions recommending that all demographers consider lagged effects when possible in light of recent research (including this study).
<!--# insert line ref -->

### 2021-08-24 doc/paper.Rmd
add more detail on limiations of DLNMs and related methods (#80)

Finally, no analytical approach assessing the potential for demographic lags can compensate for a lack of long-term data [@tenhumbergTimelaggedEffectsWeather2018; @eversLaggedDormantSeason2021].
Unfortunately, long-term data monitoring the entire life-cycle of tropical taxa are rare, and those doing so in fragmented landscapes are virtually nonexistent [@brunaCompensatoryResponsesUnderstory2005].
Without investing in collecting such data, generalizations regarding the demographic consequences of climate change in these species rich and increasingly fragmented habitats will continue to prove elusive.
More generally, however, researchers need to consider how delayed responses to climate could influence the interpretation of data in studies where the organisms lifespan exceeds the study's duration.

>>>>

Recent research, including our own presented here, shows that lagged effects of climate drivers on demography may be the norm [@eversLaggedDormantSeason2021].
As such, we suggest that all demographers investigate the possibility of lagged environmental effects, ideally using methods that do not require an *a priori* choice of lag times such as DLNMs, automated algorithms for choosing critical weather windows[@vandepolIdentifyingBestClimatic2016], or Bayesian methods that incorporate "ecological memory" [@ogleQuantifyingEcologicalMemory2015].

Finally, any analytical approach for assessing lagged effects on demography requires long-term data and/or data collected from many sites with independent weather (but similar climate and site conditions) [@tenhumbergTimelaggedEffectsWeather2018; @eversLaggedDormantSeason2021] <!--# add citation to Aldo's ESA talk -->.
Unfortunately, long-term data monitoring the entire life-cycle of tropical taxa are rare, and those doing so in fragmented landscapes are virtually nonexistent [@brunaCompensatoryResponsesUnderstory2005].
Without investing in collecting such data, generalizations regarding the demographic consequences of climate change in these species rich and increasingly fragmented habitats will continue to prove elusive.
More generally, however, researchers need to consider how delayed responses to climate could influence the interpretation of data in studies where the organisms lifespan exceeds the study's duration.
